Optimal updating factor in adaptive flat-distribution-sampling simulations

### 2017-06-13 doc/invt.tex
invt.tex: add LJ results

\documentclass[reprint, superscriptaddress, floatfix]{revtex4-1}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{upgreek}
\usepackage[table,usenames,dvipsnames]{xcolor}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{hyperref}
\usepackage{setspace}

@@

Many methods\cite{mezei1987, berg1992, *lee1993,
wang2001, *wang2001pre, huber1994,
*laio2002, *laio2008, *barducci2011, *sutto2012}
have been developed to find the proper bias potential.
%
In a class of equilibrium FDS methods\cite{
  mezei1987, berg1992, *lee1993, marinari1992, *lyubartsev1992},
the bias potential, $u_i$, is
estimated a priori and fixed
during simulation,
%
and a correction to the bias potential
is derived from the normalized histogram, $H_i$,
accumulated from the simulation as
%
%Using $H_i$ as an estimate of $\pi_i$ in
%Eq. \eqref{eq:pi_p_V}
%allows us to correct $u_i$ as
%
\begin{equation}
  u^\mathrm{corr}_i
  =
  u_i
  +
  \ln \frac{ H_i }
           { p_i }.
  \label{eq:vcorr_equil}
\end{equation}
%
\note{This follows from
  $$
  \begin{aligned}
    H_i \approx \pi_i
    &\propto p^*_i \, e^{-u_i},
    \\
    p_i
    &\propto p^*_i \, e^{-v^{0, \mathrm{corr}}_i}.
  \end{aligned}
  $$
}
However, this formula often requires a long
accumulation period for a precise histogram,
and thus disallows
continuous improvement of the bias potential
%(hence that of the PMF)
as simulation lengthens.

@@

Second, the bias potential following Eq. \eqref{eq:wl_update}
generally grows over time.
%
Since a uniform increase of $u_i(t)$, or, equivalently, that of $v_i(t)$,
does not affect the resulting distribution, $\pi_i(t)$,
the real deviation of the bias potential
depends only on the difference between $v_i(t)$
and a baseline value, $\bav{v}(t)$,
for average growth\cite{
dama2014}.
%
We define a bin-average-shifted bias potential as
%
\begin{equation}
  v_{*i}(t) \equiv v_i(t) - \bav{v}(t)
  ,
\label{eq:x_def}
\end{equation}
%
where the bin average, $\bav{v}(t)$, is defined as
%
\begin{equation}
  \bav{v}(t)
  \equiv \sum_{i = 1}^n \rho_i \, v_i(t)
  ,
\label{eq:vav_def}
\end{equation}
for some distribution,
$\pmb \rho = (\rho_1, \dots, \rho_n)$
(with $\rho_i \ge 0$ and $\sum_{i = 1}^n \rho_i = 1$).
%
Below we will use the notations, $f_{*i}$ and $\bav{f}$,
for an arbitrary variable, $f_i$.
%
Note that the distribution, $\pmb\rho$, is arbitrary
and independent of
the target distribution,
$\mathbf p = (p_1, \dots, p_n)$,
although it is usually flat.\footnote{In
well-tempered metadynamics,
$\mathbf p$ is an attenuated version of
$\mathbf p^*$\cite{
  barducci2008, dama2014}:
$\ln p_i = \ln {p^*_i}/(1+\Delta) + \mathrm{const.}$ for some positive $\Delta$,
whereas $\pmb\rho$ is flat.}

@@

%Although the above error weighting distribution, $\pmb\rho$,
%is usually flat,
%i.e. $\rho_i \equiv 1/n$,
%in practice,
%a non-flat distribution
%is sometimes more convenient in analyzing
%some multiple-bin updating schemes later on.
%
%
%From Eqs. \eqref{eq:x_def} and \eqref{eq:vav_def},
%we have
%
%\begin{equation}
%  \sum_{i = 1}^n \rho_i \, v_{*i}(t) = 0
%  .
%\label{eq:rhox_sum}
%\end{equation}
%

@@

\item \label{step:qT}
Compute the optimal $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
\footnote{We solved the equation for $\ln q(T)$ using
    a variant of the Newton-Raphson method
    with the bisection method as a fallback.
    The initial value is set to $q(T) = \ln(1+T\,a_0/2)$.}
%
%Since $m(Q)$ is a decreasing function of $Q$,
%we can start from an interval $[Q_L, Q_R]$
%that satisfies $m(Q_L) > 2/(T\,a_0) > m(Q_R)$
%(e.g. we can choose $Q_L = 0$,
%  and repeatedly double $Q_R$ until
%  the latter inequality is satisfied).
%%
%We can then iteratively refine the interval
%by substituting $Q_M = (Q_L + Q_R)/2$
%for either $Q_L$ and $Q_R$,
%depending on the sign of $m(Q_M) - 2/(T \, a_0)$,
%until the width of the interval is less than
%the desired precision,
%%
%and set $q(T)$ to $Q_M$.

\item \label{step:alpha}
Compute the optimal schedule by
  numerically integrating Eqs. \eqref{eq:q_opt} and \eqref{eq:mint}.\footnote{
  We have used the trapezoidal rule
  on a $N+1$ point logarithm grid over $[0, q(T)]$:
    $q_{[i]} = \bigl(q(T) + 1\bigr) - \bigl(q(T)+1\bigr)^{1-i/N}$,
  where $i = 0, \dots, N$, and we chose $N=20000$.
%
The integration yields a curve $q(t)$ on the grid points,
$(q_{[i]}, t_{[i]})$.
%
The updating magnitude can then be obtained from numerical differentiation:
    $\alpha_i = (1 - e^{-\lambda_1 \, \hat \alpha_i})/\lambda_1$
    and
    $\hat \alpha_i \equiv (q_{[i+1]} - q_{[i-1]}) / (t_{[i+1]} - t_{[i-1]})$,
with
$q_{[ -1]} \equiv 0$,
$q_{[N+1]} \equiv q(T)$,
$t_{[ -1]} \equiv 0$, and
$t_{[N+1]} \equiv T$.
%
The resulting values of $(t_{[i]}, \alpha_{[i]})$
can be linear interpolated
to get the updating magnitude, $\alpha(t)$, at any simulation time, $t$.
}

@@

The above procedure is still somewhat idealized.
%
The preliminary run in step \ref{step:prerun}
needs to be sufficiently long to precisely
estimate the $\Gamma_k$'s and the initial bias, $\epsilon_k$
(for which one may use the histogram-corrected bias potential).
%
However,

@@

Now that we can compute the optimal schedule
for a given updating scheme,
we will turn to the problem of finding
optimal updating schemes below.

@@

100 particle
\begin{table}[h]\footnotesize
  \caption{\label{tab:error_LJ}
    Error of the LJ system
    \note{\newline
      The data were collected in
      \texttt{data/tab1} by running
      the program \texttt{invt}.
    }%
  }
  \setlength{\tabcolsep}{2pt}
  \renewcommand\arraystretch{1.2}
  \begin{tabular} { l | p{2.2cm} p{2.4cm} p{2.0cm} }
    \hline
      XXX
    &
    &
    &
    \\
    \hline
    Global
    &
    &
    &
    \\
    \hline
    Local
    &
    &
    &
    \\
    \hline
  \end{tabular}
\end{table}

@@

\section{\label{sec:equilerr}
Inverse-time schedule as a result of increasing sample size
in a runtime average
}

The optimality of the inverse-time schedule can be considered
as a natural consequence of the stabilization of a runtime average
as the sample size increases\cite{
  marsili2006, barducci2008}.
%
We first note that the normalized histogram, $H_i(t)$,
is a runtime average of the instantaneous histogram, $h_i(t)$.
%
\begin{equation*}
  H_i(t)
  =
  \frac{ h_i(1) + \cdots + h_i(t) } { t }
  .
\end{equation*}
%
This formula can written as a recurrence relation regarding
the addition of the last data point,
%
\begin{align}
  H_i(t)
  &=
  H_i(t-1) + \frac{ h_i(t) - H_i(t-1) } { t }
  ,
  \label{eq:av_recur}
\end{align}
%
%If we interpret $x(t) - X(t-1)$
%as the independent correction to the previous average,
%$X(t-1)$, proposed by the last data point, $x(t)$,
%then this correction is absorbed into the average with
%a diminishing weight $1/t$ as the sample size, $t$, increases.
%
which allows us to incrementally improve the estimated
bias potential given by Eq. \eqref{eq:vcorr_equil}.
%
With
$u^\mathrm{corr}_i \to u_i(t+1)$, $u_i \to u_i(1)$,
we have
%
\begin{equation*}
  u_i(t+1) = u_i(1) +
  \ln \frac{ H_i(t) } { p_i }
  \approx
  u_i(1) +
  \frac{ H_i(t) } { p_i } - 1
  ,
\end{equation*}
where we have assumed that
the normalized histogram, $H_i(t)$,
is sufficiently precise in the second step.
%
%where the ``$-1$'' on the right-hand side
%can be ignored as a uniform shift
%that does not affect the distribution.
%
Thus,
%Since $H_i(t)$ is the average of the instantaneous histogram, $h_i(t)$,
%we may apply Eq. \eqref{eq:av_recur}, and
%
\begin{align}
  u_i(t+1)
  &=
  u_i(t)
  +
  \frac{ H_i(t) - H_i(t-1) }{ p_i }
  \notag
  \\
  &=
  u_i(t)
  +
  \frac{1}{t} \frac{ h_i(t) }{ p_i }
  -\frac{1}{t} \frac{ H_i(t-1) }{ p_i }
  .
  \label{eq:vbar_update}
\end{align}
%
Now if we drop the last term,
which tends to a uniform shift of the bias potential
in the asymptotic limit,
Eq. \eqref{eq:vbar_update} recovers
the single-bin updating scheme, Eq. \eqref{eq:wl_update},
under
the inverse-time schedule, Eq. \eqref{eq:alpha_invt}.

The above demonstration implies several connections
between the WL algorithm under the optimal inverse-time schedule
with a few related FDS methods.
%
First,
it establishes an efficiency equivalence
between an adaptive WL simulation under the inverse-time schedule
and the ideal equilibrium FDS counterpart
with a nearly perfect bias potential.
%
The bias potentials from the two approaches
would reach similar precision at long times.
%
Thus, it is unnecessary to convert
a sufficiently convergent adaptive FDS simulation
to the corresponding equilibrium FDS simulation
with a fixed bias potential just for better precision
of the final PMF.

Second,
it also relates a WL simulation under the inverse-time schedule
to a time-averaged variant of the regular WL simulation.
%
In the variant, the updating magnitude is fixed,
but the bias potential over the simulation course is averaged
as the final result\cite{zhou2005}.
%
The above demonstration shows that the inverse-time schedule
has already effectively implemented an implicit average
so that no explicit averaging is needed in the former approach.
%
The inverse-time schedule approach is preferred, however,
because with a decreasing updating magnitude,
it minimizes the systematic error caused by
the on-the-fly updating scheme\cite{
  zhou2005, zhou2008},
which is ignored in this analysis.

Third,
it is possible to reformulate the WL algorithm
under the inverse-time schedule
as an explicit adaptive average method
such that the bias potential is computed from
a weighted runtime average of the histogram\cite{
  marsili2006, barducci2008},
which is analogous to the adaptive bias force method\cite{
  darve2001, *darve2008}.

>>>>

\documentclass[reprint, superscriptaddress, floatfix]{revtex4-1}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{upgreek}
\usepackage[table,usenames,dvipsnames]{xcolor}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{hyperref}
\usepackage{setspace}

@@

Many methods\cite{mezei1987, berg1992, *lee1993,
wang2001, *wang2001pre, huber1994,
*laio2002, *laio2008, *barducci2011, *sutto2012}
have been developed to find the proper bias potential.
%
In a class of equilibrium FDS methods\cite{
  mezei1987, berg1992, *lee1993, marinari1992, *lyubartsev1992},
the bias potential, $u_i$, is
estimated a priori and fixed
during simulation,
%
and a correction to the bias potential
is derived from the normalized histogram, $H_i$,
accumulated from the simulation as
%
%Using $H_i$ as an estimate of $\pi_i$ in
%Eq. \eqref{eq:pi_p_V}
%allows us to correct $u_i$ as
%
\begin{equation}
  \hat u_i
  =
  u_i
  +
  \ln \frac{ H_i }
           { p_i }.
  \label{eq:vcorr_equil}
\end{equation}
%
\note{This follows from
  $$
  \begin{aligned}
    H_i \approx \pi_i
    &\propto p^*_i \, e^{-u_i},
    \\
    p_i
    &\propto p^*_i \, e^{-\hat u_i}.
  \end{aligned}
  $$
}
However, this formula often requires a long
accumulation period for a precise histogram,
and thus disallows
continuous improvement of the bias potential
%(hence that of the PMF)
as simulation lengthens.

@@

Second, the bias potential following Eq. \eqref{eq:wl_update}
generally grows over time.
%
Since a uniform increase of $u_i(t)$, or, equivalently, that of $v_i(t)$,
does not affect the resulting distribution, $\pi_i(t)$,
the real deviation of the bias potential
depends only on the difference between $v_i(t)$
and a baseline value, $\bav{v}(t)$,
for average growth\cite{
dama2014}.
%
We define a bin-average-shifted bias potential as
%
\begin{equation}
  v_{*i}(t) \equiv v_i(t) - \bav{v}(t)
  ,
\label{eq:x_def}
\end{equation}
%
where the bin average, $\bav{v}(t)$, is defined as
%
\begin{equation}
  \bav{v}(t)
  \equiv \sum_{i = 1}^n \rho_i \, v_i(t)
  ,
\label{eq:vav_def}
\end{equation}
for some distribution,
$\pmb \rho = (\rho_1, \dots, \rho_n)$
(with $\rho_i \ge 0$ and $\sum_{i = 1}^n \rho_i = 1$).
%
Below we will use the notations, $f_{*i}$ and $\bav{f}$,
for an arbitrary variable, $f_i$.
%
Note that the distribution, $\pmb\rho$,
can be independent of
the target distribution,
$\mathbf p = (p_1, \dots, p_n)$,
and chosen for convenience.\footnote{In
well-tempered metadynamics,
$\mathbf p$ is an attenuated version of
$\mathbf p^*$\cite{
  barducci2008, dama2014}:
$\ln p_i = \ln {p^*_i}/(1+\Delta) + \mathrm{const.}$ for some positive $\Delta$,
whereas $\pmb\rho$ is flat.}

@@

@@

\item \label{step:qT}
Compute the optimal value of $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
%
We implemented a variant of the Newton-Raphson method\cite{press3rd}
(for $\ln q(T)$ to avoid negative values)
with the bisection method\cite{press3rd} as a fallback.
The initial value is set to $q(T) = \ln(1+T\,a_0/2)$.
%
%Since $m(Q)$ is a decreasing function of $Q$,
%we can start from an interval $[Q_L, Q_R]$
%that satisfies $m(Q_L) > 2/(T\,a_0) > m(Q_R)$
%(e.g. we can choose $Q_L = 0$,
%  and repeatedly double $Q_R$ until
%  the latter inequality is satisfied).
%%
%We can then iteratively refine the interval
%by substituting $Q_M = (Q_L + Q_R)/2$
%for either $Q_L$ and $Q_R$,
%depending on the sign of $m(Q_M) - 2/(T \, a_0)$,
%until the width of the interval is less than
%the desired precision,
%%
%and set $q(T)$ to $Q_M$.

\item \label{step:alpha}
Compute the optimal schedule by
  numerically integrating Eqs. \eqref{eq:q_opt} and \eqref{eq:mint}.
%
To handle Gaussian updating schemes more effectively,
  we have chosen the $\mathcal N+1$ grid points over $[0, q(T)]$ to be given by
  $q_{[i]} = \bigl(q(T) + 1\bigr) - \bigl(q(T)+1\bigr)^{1-i/\mathcal N}$,
  where $i = 0, \dots, \mathcal N$, and we chose $\mathcal N=20000$.
%
The curve $q(t)$ is then integrated
  using the trapezoidal rule\cite{press3rd},
  resulting in a list of values, $\bigl(q_{[i]}, t_{[i]} \bigr)$.
%
We then compute the updating magnitude by numerical differentiation:
  $\alpha_i = (1 - e^{-\lambda_1 \, \hat \alpha_i})/\lambda_1$
  and
  $\hat \alpha_i \equiv (q_{[i+1]} - q_{[i-1]}) / (t_{[i+1]} - t_{[i-1]})$,
  with
  $q_{[ -1]} \equiv 0$,
  $q_{[N+1]} \equiv q(T)$,
  $t_{[ -1]} \equiv 0$, and
  $t_{[N+1]} \equiv T$.
%
The resulting values of $(t_{[i]}, \alpha_{[i]})$
  can be linear interpolated
  to get the updating magnitude, $\alpha(t)$, at any simulation time, $t$.

@@

The above procedure is not always necessary.
%
As shown below, for a class of bandpass updating schemes,
  which resemble the Gaussian updating schemes,
  the asymptotically optimal schedule
  is simply given by the inverse-time schedule, Eq. \eqref{eq:alpha_invt1}.
%
For these updating schemes,
  the preliminary run with a constant updating magntiude
  can be replaced by the WL-style stagewise reduction
  until the updating magnitude falls under
  the value given the inverse-time prescription\cite{
    belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}.

@@

Now that we can compute the optimal schedule
for a given updating scheme,
we will turn to the problem of finding
optimal updating schemes.

@@

Our first test system is a Lennard-Jones fluid of $N=100$ particles,
and we wish to compute the PMF along the distance of two special particles, $r$.
%
The two special particles are mutually non-interacting
but interact normally with other particles
(they have placed on an axis parallel to a side of the cubic box),
and the intrinsic distribution, $p^*$, is often referred to as
the cavity distribution function, $y(r)$.
%
We will assume reduced units in this calculation
%
The reference values were pre-calculated from a long simulation.
%
We used the Metropolis algorithm for configuration sampling\cite{frenkel}.
%
Each step consists of five MC trials of displacing a random particle,
one for one of the two special particles,
and the rest four for the other particles.
%
The MC displacement along each dimension is uniform,
with the size given by $0.1/\varrho^*$ to ensure the acceptance ratio is around 50\%,
where $\varrho^*$ is the reduced number density.

The PMF was calculated from $r = 0$ to half of the box dimension
on a grid of size $\Delta r = 0.01$ in reduced units.
%
Two densities were tried, $\varrho^* = 0.1$ and $0.8$,
and the numbers of bins were $500$ and $250$, respectively.
%
We tested the single-bin, Gaussian, and bandpass updating schemes.
%
Each simulation had an ``equilibration'' phase
of $10^7$ steps with a fixed updating magnitude, $a_0 = 10^{-4}$,
followed by a production phase
with updating magnitude given by
either the inverse-time or optimal schedule.
%
Each simulation was repeated multiple times
and the average are reported.
%
For multiple runs under the same condition,
the optimal schedule was pre-computed from a set of fixed values of
$\Gamma_k$ and $\epsilon_k$,
although they can also be fairly accurately estimated from the data
collected from the equilibration phase
of each individual run.

As shown in Table 1,
both the inverse-time and optimal schedule reduced
the error significantly.
%
The Gaussian scheme appeared to be superier
to the single-bin scheme in the $\varrho^* = 0.1$ case,
but not in the $\varrho^* = 0.8$ case,
and in the latter case,
an overly large width of the Gaussian
was counterproductive.
%

\begin{table}[h]\footnotesize
  \caption{\label{tab:error_LJ}
    Error of the LJ system.
    The numbers in parentheses are predicted values;
    $\Err_\mathrm{init.}$
    denotes the error
    at the end of the equilibration phase with a constant magnitude;
    $\Err_\mathrm{final}$
    denotes the error at the end of the production phase under
    either the inverse-time ($1/t$) or optimal schedule (opt.).
  }
  \setlength{\tabcolsep}{2pt}
  \renewcommand\arraystretch{1.4}
  \begin{tabular} { l c c c c }
    \hline
    Scheme & Sched. &
    $\Err_\mathrm{init.}$ &
    $\Err_\mathrm{final}$
    \\
    \hline
    \multicolumn{4}{c}{$\varrho^* = 0.1$, $500$ bins} \\
    \hline
    single-bin & $1/t$
    & $5.7(5.7)\times10^{-2}$
    & $9.1(11.4)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.1$ & $1/t$
    & $7.7(8.9)\times10^{-3}$
    & $6.8(8.2)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.1$ & opt.
    & $7.6\times10^{-3}$
    & $3.2\times10^{-5}$
    \\
    %\hline
    $\sigma=0.2$ & $1/t$
    & $6.5\times10^{-3}$
    & $4.0\times10^{-5}$
    \\
    %\hline
    $\sigma=0.2$ & opt.
    & $7.0\times10^{-3}$
    & $2.5\times10^{-5}$
    \\
    %\hline
    $\sigma=0.5$ & $1/t$
    & $5.5\times10^{-3}$
    & $3.2\times10^{-5}$
    \\
    %\hline
    $\sigma=0.5$ & opt.
    & $5.8\times10^{-3}$
    & $2.6\times10^{-5}$
    \\
    %\hline
    $K=20$ & $1/t$
    & $8.1(9.0)\times10^{-3}$
    & $1.6(1.8)\times10^{-5}$
    \\
    \hline
    \multicolumn{4}{c}{$\varrho^* = 0.8$, $250$ bins} \\
    \hline
    single-bin & $1/t$
    & $1.3(1.4)$
    & $2.7(2.7)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.1$ & $1/t$
    & $1.3(1.3)$
    & $2.9(2.8)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.1$ & opt.
    & $1.2$
    & $2.8\times10^{-3}$
    \\
    %\hline
    $\sigma=0.2$ & $1/t$
    & $1.3$
    & $3.4\times10^{-3}$
    \\
    %\hline
    $\sigma=0.2$ & opt.
    & $1.2$
    & $3.1\times10^{-3}$
    \\
    %\hline
    $\sigma=0.5$ & $1/t$
    & $0.95$
    & $9.8\times10^{-3}$
    \\
    %\hline
    $\sigma=0.5$ & opt.
    & $1.0$
    & $5.8\times10^{-3}$
    \\
    %\hline
    $K=20$ & $1/t$
    & $1.3(1.3)$
    & $3.0(2.7)\times10^{-3}$
    \\
    \hline
  \end{tabular}
\end{table}

@@

\section{\label{sec:equilerr}
Runtime average interpretation
of the inverse-time schedule
}

The optimality of the inverse-time schedule can be considered
as a natural consequence of the stabilization of a runtime average
as the sample size increases\cite{
  marsili2006, barducci2008}.
%
We first note that a runtime average
\begin{equation}
  X(t) = \frac{x(1) + \cdots + x(t)}{t}
  .
  \label{eq:avx}
\end{equation}
can be written as a recurrence relation
\begin{equation}
  X(t) = X(t - 1) + \frac{x(t) - X(t-1)}{t}
  .
  \label{eq:avx_recur}
\end{equation}
%
Here $x(t) - X(t-1)$ serves as an independent correction
from data point at time $t$ to the previous average value,
and this correction is absorbed into the new value, $X(t)$,
with a weight of $1/t$,
which is the inverse sample size.

We can use Eq. \eqref{eq:avx_recur} to incrementally
improve the estimated
bias potential given by Eq. \eqref{eq:vcorr_equil}.
%
Assuming the correction, $\hat u_i$, is done at every step,
then
%
\begin{equation*}
  \hat u_i(t+1) = u_i(1) +
  \ln \frac{ H_i(t) } { p_i }
  \approx
  u_i(1) +
  \frac{ H_i(t) } { p_i } - 1
  ,
\end{equation*}
where we have assumed that
the normalized histogram, $H_i(t)$,
is sufficiently precise in the second step.
%
Since the normalized histogram, $H_i(t)$,
is an average of the instantaneous histogram, $h_i(t)$,
$H_i(t) = \frac 1 t \sum_{\tau = 1}^t h_i(\tau)$,
we have, from Eq. \eqref{eq:avx_recur},
%
\begin{align}
  H_i(t)
  &=
  H_i(t-1) + \frac{ h_i(t) - H_i(t-1) } { t }
  ,
  \label{eq:avH_recur}
\end{align}
%
Thus,
%Since $H_i(t)$ is the average of the instantaneous histogram, $h_i(t)$,
%we may apply Eq. \eqref{eq:av_recur}, and
%
\begin{align}
  \hat u_i(t+1)
  &=
  \hat u_i(t)
  +
  \frac{ H_i(t) - H_i(t-1) }{ p_i }
  \notag
  \\
  &=
  \hat u_i(t)
  +
  \frac{1}{t} \frac{ h_i(t) }{ p_i }
  -\frac{1}{t} \frac{ H_i(t-1) }{ p_i }
  .
  \label{eq:vhat_update}
\end{align}
%
Now if we drop the last term,
which tends to a uniform shift of the bias potential
in the asymptotic limit,
Eq. \eqref{eq:vhat_update} recovers
the single-bin updating scheme, Eq. \eqref{eq:wl_update},
under
the inverse-time schedule, Eq. \eqref{eq:alpha_invt}.

The above demonstration implies several connections
between the WL algorithm under the optimal inverse-time schedule
with a few related FDS methods.
%
First,
it establishes an efficiency equivalence
between an adaptive WL simulation under the inverse-time schedule
and one with a fixed and nearly perfect bias potential.
%
Since the bias potential in the former approach
essentially follows the same time evolution
for the correction,
$\hat u_i(t)$, Eq. \eqref{eq:vhat_update},
the two would reach similar precision at long times.
%
Thus, it is unnecessary to fix the bias potential
after an adaptive FDS simulation has reached
sufficiently convergent just for better precision
of the final PMF.

Second,
it shows close connection between
the WL algorithm under the inverse-time schedule
and self-healing umbrella sampling.
In the latter,
the bias potential is computed from
a weighted runtime average of the histogram\cite{
  marsili2006, barducci2008},
which is analogous to the adaptive bias force method\cite{
  darve2001, *darve2008}.

Third,
it also relates a WL simulation under the inverse-time schedule
to a time-averaged variant of the regular WL simulation.
%
In the variant, the updating magnitude is fixed,
but the bias potential over the simulation course is averaged
as the final result\cite{zhou2005}.
%
The above demonstration shows that the inverse-time schedule
has already effectively implemented an implicit average
so that no explicit averaging is needed in the former approach.
%
The inverse-time schedule approach is preferred, however,
because with a decreasing updating magnitude,
it minimizes the systematic error caused by
the on-the-fly updating scheme\cite{
  zhou2005, zhou2008},
which is ignored in this analysis.
